---
title: "R Notebook"
output: html_notebook
---
<style>
th, td {
    padding-left: 3%;
}

</style>



```{r message=FALSE, include=FALSE, echo=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(stargazer)
library(tibble)
library(psych)
```

```{r include=FALSE, echo=FALSE}
crime_85_17_raw <- read.csv('./data/crime/crime_clearances_85-17.csv', stringsAsFactors = FALSE)
#pull in evictions data
evictions_raw <- read.csv("./data/evictions/counties-eviction-lab.csv")
```


# Cleaning the Crime Data


*****

```{r}
# filter data to 2000 to 2016
crime_00_16_raw <- crime_85_17_raw %>% filter(Year %in% c(2000:2016))

# select only violent crimes (homicide, rape, robbery, and aggravated assault)
crime_00_16_violent <- crime_00_16_raw %>% select(Year:NCICCode, Violent_sum, Homicide_sum, ForRape_sum, Robbery_sum, AggAssault_sum)

crime_00_16_violent <- crime_00_16_violent %>% group_by(County, Year) %>% summarise(violent_crime = sum(Violent_sum), homicide = sum(Homicide_sum), rape = sum(ForRape_sum), robbery = sum(Robbery_sum), agg_assault = sum(AggAssault_sum))


#view the crime data
head(crime_00_16_violent) %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center") %>%
  scroll_box(width = "100%", height = "15%")
```


# Cleaning the Evictions Data (for Merging with Crimes Data)

```{r}
# create unique id for evictions data
evictions_name <- evictions_raw %>% mutate(name_no_space = gsub(" ","_",name))
evictions <- evictions_name %>% mutate(uid = tolower(paste(as.character(name_no_space), as.character(year), sep="_")))
```

```{r}
# create unique id for crimes data and merge column
crime_name <- crime_00_16_violent %>% mutate(name_no_space = gsub(" ","_",County))
crime <- crime_name %>% mutate(uid = tolower(paste(as.character(name_no_space), as.character(Year), sep="_")))
```

```{r}
# join the two datasets
crime_evictions_raw <- left_join(crime,evictions, by="uid")
```

```{r}
# The joined datasets 
data <- crime_evictions_raw %>% select(-c(year, name, name_no_space.y, name_no_space.x, parent.location)) %>% select(County:agg_assault, uid:subbed)
 
head(data) %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center") %>%
  scroll_box(width = "100%")
```




```{r}

data %>% filter(violent_crime == max(violent_crime) & population == min(population)) %>% group_by(County) %>% filter()
```

```{r}
data %>% group_by(County) %>% summarise(vcp = sum(violent_crime_p)) %>% filter(vcp == min(vcp))
```


The County with an average violent crime count above the 75th quantile ( i.e. the countioes with the most violent crime), but the smallest population (in that group)

```{r}
data
data_sum <- data %>% group_by(County) %>% summarise(vc_sum = mean(violent_crime), pop = mean(population), vcpop = mean(violent_crime_p)) 
quantile(data_sum$vc_sum)
quantile(data_sum$pop)

most_county_name <- data_sum %>% filter(vc_sum >= 2192) %>% arrange(pop) %>% filter(pop == min(pop)) %>% select(County)
most_county_name <- most_county_name$County


data_sum %>% arrange(desc(vc_sum))
```

The County with an average violent crime count above the 25th quantile ( i.e. the counties with the least violent crime), but the smallest population (in that group)

```{r}
least_county_name <- data_sum %>% filter(vc_sum <= 141) %>% arrange(desc(pop)) %>% filter(pop == max(pop)) %>% select(County)
#least_county_name <- least_county_name$County
least_county_name <- "Calaveras County"
```



# Analysis 


Making crime counts proportional to population size:

```{r}
data <- data %>% mutate(violent_crime_p = violent_crime/population, homicide_p = homicide/population, rape_p = rape/population, robbery_p = robbery/population, agg_assault_p = agg_assault/population)
```

Making # evictions filed  proportional to population size:

```{r}
data <- data %>% mutate(eviction_filings_p = eviction.filings/population)
```


```{r}
head(data) %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center") %>%
  scroll_box(width = "100%")
```



County with Least Violence But Large Population 
vs
County with A lot fo violence but small population
```{r}
least_county_name
most_county_name
```


```{r}
main_counties <- data %>% filter(County == least_county_name | County == most_county_name) 
not_main_counties <- data %>% filter(County != least_county_name & County != most_county_name)


main_counties %>% filter(County == most_county_name) 
```






```{r}
ggplot(main_counties, aes(Year, violent_crime, color=County))+
  geom_line()+
  #geom_line(aes(Year, eviction_filings_p), linetype="longdash")+
  #geom_point(aes(Year, eviction_filings_p),fill=pct.white)+
  labs(title="Violent Crime in Tulare and Calaveras County",
       subtitle="2000 to 2016",
       caption="Data Source: Open Justice", x="Year", y="Violent Crime Count") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "bottom") +
  ggsave("violent_crime_2counties.png")
```


```{r}
ggplot(data_sum %>% filter(County != "Los Angeles County"), aes(vc_sum, pop, color=County))+
  geom_point()+
  geom_hline(yintercept = mean(data_sum$pop)) +
  #geom_line(aes(Year, eviction_filings_p), linetype="longdash")+
  #geom_point(aes(Year, eviction_filings_p),fill=pct.white)+
  # labs(title="Violent Crime in Tulare and Calaveras County",
  #      subtitle="2000 to 2016",
  #      caption="Data Source: Open Justice", x="Year", y="Violent Crime Count") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "none")
```

```{r}
data_sum %>% filter(County == "Stanislaus County" | County == "Ventura County") 

data_sum %>% mutate(vcpop100 = vcpop*100) %>% arrange(desc(vcpop100), pop, desc(vc_sum))
```

```{r}
most_v_counties <- data %>% group_by(County) %>% summarise(vc = mean(violent_crime), vcp = mean(violent_crime_p)) %>% filter(vcp > mean(vcp)) %>% arrange(desc(vc)) %>% select(County) %>% head(5)

most_v <- data %>% filter(County %in% most_v_counties$County)
```



```{r}
ggplot(most_v, aes(Year, violent_crime, color=County))+
  geom_line()+
  #geom_line(aes(Year, evictions))+
  #geom_line(aes(Year, eviction_filings_p), linetype="longdash")+
  #geom_point(aes(Year, eviction_filings_p),fill=pct.white)+
  labs(title="Violent Crime in Tulare and Calaveras County",
       subtitle="2000 to 2016",
       caption="Data Source: Open Justice", x="Year", y="Violent Crime Count") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))
  #ggsave("violent_crime_2counties.png")
```




*******


```{r}
0.8653375-0.5873318
```




The population in SJ county is less than Fresno County, however, the Crime rate (total crime/population) is higher. 

```{r}
sj_fr <- data %>% filter(County == "San Francisco County" | County == "Alpine County")
```



```{r}
ggplot(sj_fr, aes(Year, violent_crime, color=County))+
  geom_line()+
  geom_line(aes(Year, evictions), linetype="dotdash")+
  #geom_line(aes(Year, eviction_filings_p), linetype="longdash")+
  #geom_point(aes(Year, eviction_filings_p),fill=pct.white)+
  labs(title="Violent Crime in San Joaquin and Fresno County",
       subtitle="2000 to 2016",
       caption="Data Source: Open Justice", x="Year", y="Violent Crime Count") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "bottom")
  #ggsave("violent_crime_2counties.png")
```



```{r}

```




```{r}
data %>% group_by(County) %>% summarise(vc = sum(violent_crime)) %>% arrange((vc))
data %>% filter(County== "Alpine County")
```


```{r}
data <- data %>% mutate(
  non_white = ifelse(pct.white < 1, 1, 0)
  )
```


```{r}
ggplot(data %>% filter(County == "Los Angeles County" | County == "San Joaquin County"), aes(Year, violent_crime, color=County))+
  geom_line()+
  geom_line(aes(Year, evictions), linetype="dotdash")+
  #geom_line(aes(Year, eviction_filings_p), linetype="longdash")+
  #geom_point(aes(Year, eviction_filings_p),fill=pct.white)+
  labs(title="Violent Crime in San Joaquin and Fresno County",
       subtitle="2000 to 2016",
       caption="Data Source: Open Justice", x="Year", y="Violent Crime Count") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "bottom")
  #ggsave("violent_crime_2counties.png")
```



```{r}
ggplot(data, aes(County, evictions))+
  geom_col(aes(fill=pct.white))+
  #geom_line(aes(Year, eviction_filings_p), linetype="longdash")+
  #geom_point(aes(Year, eviction_filings_p),fill=pct.white)+
  labs(title="Violent Crime in San Joaquin and Fresno County",
       subtitle="2000 to 2016",
       caption="Data Source: Open Justice", x="Year", y="Violent Crime Count") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "bottom")
  #ggsave("violent_crime_2counties.png")
```

```{r}
inst <- data %>% group_by(County) %>% summarise(ev = sum(evictions), evf = sum(eviction.filings), vc = sum(violent_crime)) %>% arrange(desc(evf)) %>% filter(!is.na(evf)) %>% filter(County != "Alpine County") %>% select(County)

high_names <- inst[2:4,]
mid_names <- inst[10:12,]
low_names <- inst[21:23,]

high_e <- data %>% filter(County %in% high_names$County) %>% mutate(instability_level = "Most Unstable")
med_e <- data %>% filter(County %in% mid_names$County) %>% mutate(instability_level = "Median Unstable")
low_e <- data %>% filter(County %in% low_names$County) %>% mutate(instability_level = "Least Unstable")

high_v <- data %>% filter(County %in% high_names$County) %>% mutate(violence_level = "Most Violent")
med_v <- data %>% filter(County %in% mid_names$County) %>% mutate(violence_level = "Median Violent")
low_v <- data %>% filter(County %in% low_names$County) %>% mutate(violence_level = "Least Violent")


hml_e <- rbind(high_e,med_e,low_e)
hml_v <- rbind(high_v,med_v,low_v)
```

```{r}
ggplot(hml_e, aes(County, violent_crime))+
  geom_col(aes(fill=instability_level))+
  #geom_line(aes(Year, eviction_filings_p), linetype="longdash")+
  #geom_point(aes(Year, eviction_filings_p),fill=pct.white)+
  labs(title="Violent Crime Counts by Eviction Filing Level in CA",
       subtitle="2000 to 2016",
       caption="Data Source: Open Justice & Eviction Lab", x="County", y="Violent Crime Count") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1, size=9))+
  ggsave("./final_graphs/violent_crime_2counties.png")

```


```{r}
quantile(data$eviction.filings, na.rm=TRUE)
data <- data %>% mutate(
  instability_level = ifelse(eviction.filings > 1700, "Most Unstable", ifelse(eviction.filings <= 1700 & eviction.filings > 85, "Median Unstable", ifelse(eviction.filings <= 85, "Least Stable", 0)))
)
```


```{r}
model_1 <- lm(violent_crime ~ factor(instability_level), data=data)
stargazer(model_1, type="text")
```


```{r}
levs <- c("Least Instability", "Median Instability", "Most Instability")
amts <- c(76.214, 939.570, 12069.650)
sigs <- c("...", "...", "p<.01")

effect_unst <- data.frame(levs = levs, amts = amts, Significance = sigs)

ggplot(effect_unst, aes(levs, amts, shape=Significance, color=Significance))+
  geom_point(size=5)+
  labs(title="Effect of Eviction Filings Level on Violent Crime Count",
       subtitle="Aggregated Data from 2000 to 2016",
       caption="Data Source: Open Justice & Eviction Lab", x="Instability Level (Eviction Filings)", y="Average Effect Size") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))+
  ggsave("./final_graphs/effect_instability.png")
```


```{r}
# More volatility in violence, in unstable group 

ggplot(hml_e, aes(Year, violent_crime, color=instability_level, linetype=County))+
  geom_line()+
  facet_grid(cols = vars(instability_level))+ 
  labs(title="Time Trend of Violent Crime Count by Instability Level",
       subtitle="Aggregated Data from 2000 to 2016",
       caption="Data Source: Open Justice & Eviction Lab", x="Year", y="Violent Crime Count") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))+
  ggsave("./final_graphs/time_trend_violence_main.png")
```

```{r}

# Also More Volatility in Eviction Filings 

ggplot(hml_e %>% filter(instability_level=="Most Unstable"), aes(Year, eviction.filings, color=County))+
  geom_line()+
  labs(title="Time Trend of Eviction Filings for Counties with most Eviction Filings",
       subtitle="Aggregated Data from 2000 to 2016",
       caption="Data Source: Open Justice & Eviction Lab", x="Year", y="Eviction Filings") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))+
  ggsave("./final_graphs/time_trend_evictions_most_unstable.png")
```

```{r}
data_no_na <- data %>% filter(!is.na(eviction.filings))
cor(data_no_na$violent_crime, data_no_na$eviction.filings)
```
```{r}
data_year <- data_no_na %>% group_by(Year) %>% summarise(vc = sum(mean(violent_crime)), evf = sum(mean(eviction.filings)))

ggplot(data_year, aes(Year, vc))+
  geom_point(aes(color="Violent Crime Counts"))+
  geom_line(aes(color="Violent Crime Counts"))+
  geom_point(aes(Year, evf, color="Eviction Filings"))+
  geom_line(aes(Year, evf, color="Eviction Filings"))+
  labs(title="Time Trend of Eviction Filings and Violent Crime, California",
       subtitle="Aggregated Data from 2000 to 2016",
       caption="Data Source: Open Justice & Eviction Lab", x="Year", y="Violent Crime / Eviction Filings") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))+
  ggsave("./final_graphs/time_trend_total.png")
```

```{r}
data_corr <- data_no_na %>% group_by(Year) %>% summarise(corr_e_v = cor(violent_crime, eviction.filings))

ggplot(data_corr, aes(Year, corr_e_v))+
  geom_point()+
  geom_line()+
  geom_hline(yintercept = mean(data_corr$corr_e_v), color="tomato2")+
  annotate("text", x=2002, y=.97, label="Average Correlation")+
  labs(title="Correlation between Violent Crime and Eviction Filings, by Year in California",
       subtitle="Aggregated Data from 2000 to 2016",
       caption="Data Source: Open Justice & Eviction Lab", x="Year", y="Correlation") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))+
  ggsave("./final_graphs/corr_yearly.png")
```



```{r}
model_2 <- lm(violent_crime_p ~ eviction_filings_p, data = data)
stargazer(model_2, type="latex", title="Effect of Eviction Filings Rate on Violent Crime Rate")
```


$$
\frac{Eviction \space Filings}{Population}
$$
$$
\frac{Violent \space Crimes}{Population}
$$
